### 2014-04-14 paper/tubes.tex
working through stuff, adding discussion of powder-based printers

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with hollow pipes and cavities on additive manufactuing machines. Pipes will frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
%Beyond design, the tubes must actually be printable to be useful.  
%When geometries contain overhangs, as there may be in designed pipes and cavities, 
To allow overhangs, 3D printers lay down support material that must be later removed from the model.  This removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We will discuss techniques for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}}, and multi-jetting machines such as an Objet\footnote{\url{http://stratasys.com}}.

@@

\subsection{Multi-Jetting}
On multi-jetting machines, bridging is not possible and stepover potential is much less, because the machine is laying down droplets of liquid that are later UV cured.  Unsupported overhangs can be approximately $14^{\circ}$ on our Objet printer.  Multi-jetting and higher-end FDM machines use a secondary support material that must be manually removed afterwards. Some support can be melted out (Projet wax material); other support can be dissolved in lye or blasted with high-pressure water. Complex pipe geometries can make water blasting impossible and other processes more labor- and time-intensive.  All models in Figure \ref{fig:teaser}b required support material removal post-print: the neon sign and maze were additionally cut into two pieces to give better access to the support.
\bjoern{should we say something about stereolithography printers like the Form 1? we don't have one, but they could be ideal for our techniques since the liquid will just flow out of our tubes. However, they also generate support structure and I'm not sure how those supports differ from FDM.}

>>>>

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with hollow pipes and cavities on additive manufactuing machines. Pipes will frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
%Beyond design, the tubes must actually be printable to be useful.  
%When geometries contain overhangs, as there may be in designed pipes and cavities, 
To allow overhangs, 3D printers lay down support material that must be later removed from the model.  This removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We will discuss techniques for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}}, multi-jetting machines such as an Objet\footnote{\url{http://stratasys.com}}, and powder-based machines such as the ZCorp\footnote{\url{http://3dsystems.com}}.  Other varities of machines, e.g., stereolithography machines such as the Form 1\footnote{\url{http://formlabs.com}}, require similar techniques to these.

@@

\subsection{Multi-Jetting}
On multi-jetting machines, bridging is not possible and stepover potential is much less, because the machine is laying down droplets of liquid that are later UV cured.  Unsupported overhangs can be approximately $14^{\circ}$ on our Objet printer.  Multi-jetting and higher-end FDM machines use a secondary support material that must be manually removed afterwards. Some support can be melted out (Projet wax material); other support can be dissolved in lye or blasted with high-pressure water. Complex pipe geometries can make water blasting impossible and other processes more labor- and time-intensive.  All models in Figure \ref{fig:teaser}b required support material removal post-print: the neon sign and maze were additionally cut into two pieces to give better access to the support.

\subsection{Powder-based}
Powder-based machines have a bed of powder which is selectively fused together using glue droplets.  As each layer is completed, the bed moves down and another layer of powder is added on top.  The unfused powder functions as the support structure for the fused powder: this makes the support trivial to remove post-print, as the loose grains will simply fall out of pipes and cavities in the model.  These machines unfortunately cannot fabricate flexible or transparent materials (yet), however they can print in full-color.  We did not have access to a powder-based printer, though in future work we would like to experiment with one.
